\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-06-15 archive.tex
more learning from scratch

\newcommand{\greenCode}[1]{{\footnotesize\popp{\code{#1}}}}

@@

\subsection{Learning from Scratch}
A long-standing dream within the program induction community
is ``learning from scratch'': starting with a \emph{minimal} Turing-complete programming language,
and then learning to solve a wide swath of
induction problems~\cite{solomonoff1964formal}.
All existing systems,
including ours,
fall far short of this dream,
and it is unclear (and we believe unlikely)
that this dream could ever be fully realized.
How far can we push in this direction?
``Learning from scratch'' is subjective, but a reasonable
starting point is the set of primitives provided in 1959
Lisp~\cite{mccarthy1960recursive}: these include
conditionals, recursion, arithmetic, and the 
list operators \code{cons}, \code{car}, \code{cdr}, and \code{nil}.
A  basic first goal is to start with
these primitives,
and then recover a DSL that
more closely resembles modern functional languages like Haskell and OCaml.
Recall (Sec.~\ref{sequences})
that we initially provided our system with functional programming routines like
\code{map} and \code{fold}.

@@

\begin{figure}  \newcommand{\helpSize}{0.25cm}
  \begin{tabular}{cc}\toprule
    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}\\\midrule
    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
    \begin{tabular}{l}
      \greenCode{$f_0($x$)\,=\,$(+ x 1)}\\
      \hspace{\helpSize}($f_0$: \emph{increment})\\
      \greenCode{$f_1($x$)\,=\,$(- x 1)}\\
      \hspace{\helpSize}($f_1$: \emph{decrement})\\
        \popp{\code{$f_2($p$,$f$,$n$,$x$)\,=\,$(if (p x) nil}}\\
      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(cons (f x) ($f_2$ (n x))))}}\\
        \hspace{\helpSize}($f_2$: \emph{unfold})\\
        \popp{\code{$f_3($i$,$l$)\,=\,$(if (= i 0) (car l)}}\\
      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{($f_3$ ($f_1$ i) (cdr l))))}}\\
        \hspace{\helpSize}($f_3$: \emph{index})\\
        \popp{\code{$f_4($f$,$l$,$x$)\,=\,$(if (empty? l) x}}\\
      \phantom{\code{$f_4($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(f (car l) ($f_4$ (cdr l))))}}\\
        \hspace{\helpSize}($f_4$: \emph{fold})\\

>>>>

\newcommand{\greenCode}[1]{{\footnotesize\popp{\code{#1}}}}
\newcommand{\blueCode}[1]{{\footnotesize\pop{\code{#1}}}}

@@

\subsection{Learning from Scratch}
A long-standing dream within the program induction community
is ``learning from scratch'': starting with a \emph{minimal} Turing-complete programming language,
and then learning to solve a wide swath of
induction problems~\cite{solomonoff1964formal,schmidhuber2004optimal,hutter2004universal,solomonoff1989system}.
All existing systems,
including ours,
fall far short of this dream,
and it is unclear (and we believe unlikely)
that this dream could ever be fully realized.
How far can we push in this direction?
``Learning from scratch'' is subjective, but a reasonable
starting point is the set of primitives provided in 1959
Lisp~\cite{mccarthy1960recursive}: these include
conditionals, recursion, arithmetic, and the 
list operators \code{cons}, \code{car}, \code{cdr}, and \code{nil}.
A  basic first goal is to start with
these primitives,
and then recover a DSL that
more closely resembles modern functional languages like Haskell and OCaml.
Recall (Sec.~\ref{sequences})
that we initially provided our system with functional programming routines like
\code{map} and \code{fold}.

@@

\begin{figure}  \newcommand{\helpSize}{0.25cm}
  \begin{tabular}{cl}\toprule
    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&
    \begin{tabular}{l}
      \code{[4\, 7\, 6\, 9\, 0]$\to$5}\\
\code{[3\, 8\, 2\, 1\, 4]$\to$4}\\
\code{[2\, 2\, 9\, 4\, 8]$\to$3}\\
\code{[0\, 9\, 8\, 9\, 9\, 8]$\to$1}\\
\blueCode{$f(\ell) = $($f_0$ (car $\ell$))}\\\\

\code{[2\, 5\, 6\, 0\, 6]$\to$19}\\
\code{[9\, 2\, 7\, 6\, 3]$\to$27}\\
\blueCode{$f(\ell) = $($f_4$ + $\ell$ 0)}\\\\

\code{[4\, 2\, 6\, 4]$\to$[8\, 4\, 12\, 8]}\\
\code{[0\, 3\, 5\, 0\, 3\, 2]$\to$[0\, 6\, 10\, 0\, 6\, 4]}\\
\code{[2\, 3\, 0\, 7]$\to$[4\, 6\, 0\, 14]}\\
\blueCode{$f(\ell) = $($f_5$ ($\lambda$ (x) (+ x x)) $\ell$)}\\\\

      \code{[5\, 2\, 9]}$\to$\code{[[2\, 9]\, [9]\, []]}\\
      \code{[3\, 8\, 1\, 3]}$\to$\code{[[8\, 1\, 3]\, [1\, 3]\, [3]\, []]}\\
      \blueCode{$f(\ell) = $($f_2$ empty? cdr cdr $\ell$)}
      \\

      \end{tabular}
    \\\midrule
    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
    \begin{tabular}{l}
      \greenCode{$f_0($x$)\,=\,$(+ x 1)}\\
      \hspace{\helpSize}($f_0$: \emph{increment})\\
      \greenCode{$f_1($x$)\,=\,$(- x 1)}\\
      \hspace{\helpSize}($f_1$: \emph{decrement})\\
        \popp{\code{$f_2($p$,$f$,$n$,$x$)\,=\,$(if (p x) nil}}\\
      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(cons (f x) ($f_2$ (n x))))}}\\
        \hspace{\helpSize}($f_2$: \emph{unfold})\\
        \popp{\code{$f_3($i$,$l$)\,=\,$(if (= i 0) (car l)}}\\
      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{($f_3$ ($f_1$ i) (cdr l))))}}\\
        \hspace{\helpSize}($f_3$: \emph{index})\\
        \popp{\code{$f_4($f$,$l$,$x$)\,=\,$(if (empty? l) x}}\\
      \phantom{\code{$f_4($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(f (car l) ($f_4$ (cdr l))))}}\\
        \hspace{\helpSize}($f_4$: \emph{fold})\\
        \greenCode{$f_5($f$,$l$)\,=\,$(if (empty? l) nil)}\\
        \phantom{\greenCode{$f_5($f$,$l$)\,=\,$(if }}\greenCode{(cons (f (car l)) ($f_5$ (cdr l)))}\\
        \hspace{\helpSize}($f_5$: \emph{map})

%        -0.222777	int -> int -> list(int)	#(lambda (lambda (fix1 $0 (lambda (lambda (#(lambda (lambda (lambda (if $0 empty (cons $1 $2))))) (map (lambda (#(+ 1) $0)) ($1 (#(lambda (- $0 1)) $0))) 0 (eq? $0 $3)))))))
